The 4th Workshop on Sustainable Software: Best Practices and Experiences (WSSSPE4)

### 2016-06-10 WSSSPE4_proposal/WSSSPE4_proposal.tex
updates throughout

\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{enumitem}

@@

{\bf \noindent Summary:} The WSSSPE4 workshop is part of a community driven effort to address the many new challenges that are evident for the development, deployment, maintenance, and sustainabiity of reusable software for academic research and education. These challenges include the software development process, the support and maintenance of software, governance and business models particularly for sustainability, the role of software in building science communities, the need for software to lead to science which is verifiable and reproducible, policy issues such as how to measure software impact, choose licensing, or assign software credit, and the education and nurturing of early stage researchers in scientific software. 
Based on successful WSSSPE1-WSSSPE3 meetings at SC13, SC14, and NCAR in 2013, 2014, and 2015, the organizers expect over 60 submitted papers and around 75 attendees at the WSSSPE4 meeting that will be held immediately preceding the RSE conference in Manchester, UK, in September 2016.  As with previous events, WSSSPE4 will provide an open and dynamic forum for the community to discuss experiences and approaches in the areas listed above. This proposal requests funds to support attendance at WSSSPE4 for US based researchers, particularly targeted at encouraging attendance by early stage researchers with attention to diversity.

@@

This is the fourth major event in the WSSSPE series. The first two WSSSPE workshops were held as part of SC in 2013 and 2014. These one-day workshops brought together researchers and practitioners to discuss challenges and issues around the development, deployment and maintenance of reusable software for academic research and education, and discussed potential solutions, moving from a presentation to a presentation and discussion format.  WSSSPE3, held at NCAR in 2015, grew out of the previous events, and focused almost entirely on facilitated discussions, with the aim of starting working groups to make changes in the practice of scientific software development leading to more sustainable software.  Two of these groups have continued
working together, leading to the FORCE11 Software Citation Principles and an in-progress paper
on best practices for sustainable scientific software development.

@@

{\bf Track 1 – Building a sustainable future for open-use research software} has the goals of defining a vision of the future of open-use research software, and in the workshop, initiating the activities that are needed to get there.  The idea of this track is to first think about where we want to be 5 to 10 years from now, without being too concerned with where we are today, and then to determine how we can move to this future.

{\bf Track 2 – Practices \& experiences in sustainable scientific software} has the goal of improving the quality of today’s research software and the experiences of its developers by sharing practices and experiences.  This track is focused on the current state of scientific software and what we can do to improve it in the short term, starting with where we are today.

@@

The WSSSPE~2 organizing committee comprises 
\begin{itemize}
\item Co-chairs: Gabrielle Allen (Professor, University of Illinois), Daniel S. Katz (Researcher, University of Chicago)
\item From WSSSPE~1 organizing committee: Manish Parashar (Professor, Rutgers University), Neil Chue Hong (Director, University of Edinburgh), David Proctor (Currently AAAS Fellow, NSF). 
\item New organizing members for WSSSPE 2: Colin Venters (Researcher, University of Leads), Nancy Wilkins-Diehr (Researcher, UCSD), and  Karen Cranston (National Evolutionary Synthesis Center)
\end{itemize}

The organizing committee will in the next months be discussing changes to the program based on last year’s experience, but it is anticipated that the one-day workshop will include
\begin{itemize}
\item Two invited keynote speakers
\item Peer reviewed short papers that will form the basis for discussions or panels
\item Targeted panel discussions
\item Breakout groups to discuss on feed back on identified issues
\end{itemize}

The WSSSPE 1 workshop used some novel features that are described in the arXiv report (http://arxiv.org/format/1311.3523v2 and Appendix) and will probably be repeated for WSSSPE 2. These included
\begin{itemize}
\item Authors submitted papers by placing them on an open archive of their choice and submitting the URL for the document
\item Instead of individual author presentations, the accepted papers where grouped into three panels, with panelists presenting the findings from a set of papers followed by a general discussion with all participants.
\end{itemize}

The tentative agenda for the Workshop is

\begin{tabular}{|l|l|} \hline
9.00 --- 9.15 & Introduction, review of the day's activities \\ \hline
9.15 --- 10.15 & Keynotes \\ \hline
10.15 --- 10.30 & Break \\ \hline
10.30 --- 11.15 & Panel 1 and Audience Discussion \\ \hline
11.15 --- 12.00 & Panel 2 and Audience Discussion \\ \hline
12.00 --- 1.00 & Lunch \\ \hline
1.00 --- 2.30 & Led breakout sessions \\ \hline
2.30 --- 3.30 & Summaries and discussion from breakout discussions \\ \hline
3.30 --- 4.30 & Panel 3 and Audience Discussion \\ \hline
4.30 --- 5.00 & Organize follow on activities  \\ \hline
\end{tabular}

@@

We are seeking funding early enough to be able to include a description of the opportunity for travel support in our call for papers and outreach material for the event to make sure that lack of funding is not a barrier to submitting a paper, and to make it clear that participation by early stage researchers is welcomed and encouraged.

@@

The WSSSPE 2 organizing committee will review the applications and select the participants to be funded through the award.

@@

The National Center for Supercomputing Applications (NCSA) at the University of Illinois (U of I) will be the lead fiscal institution for this proposal to the Office of Advanced Cyberinfrastructure (ACI) at the National Science Foundation. NCSA staff will manage the travel for the participants and presenters funded through this proposal. US participants and presenters will arrange their flights through the travel agency of NCSA/University of Illinois, with reimbursed by NCSA/University of Illinois. NCSA/ University of Illinois will cover the hotel and meals for all US and European participants and presenters. The WSSSPE~2 organizing committee will review the applications and select the participants to be funded through this award.

@@

The focus of the WSSSPE~2 workshop on sustainable software for science and engineering is a topic of very broad relevance and interest that cuts across many different disciplines and different categories of people from students, developers, faculty, etc. Areas of interest and focus in the workshop include education and career paths of those involved in the development of software.

@@

@@

\section*{Appendix}

The following is a report on the WSSSPE~1 workshop that is available at the arXiv (http://arxiv.org/abs/1311.3523)

@@

{\bf WSSSPE2}

{\bf Principal Investigator: Gabrielle Allen}

>>>>

\documentclass[11pt]{article}
\usepackage{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{epstopdf}
\usepackage{enumitem}
\usepackage{url}
\usepackage{color}

@@

{\bf \noindent Summary:} The WSSSPE4 workshop is part of a community driven effort to address the many new challenges that are evident for the development, deployment, maintenance, and sustainability of reusable software for academic research and education. These challenges include the software development process, the support and maintenance of software, governance and business models particularly for sustainability, the role of software in building science communities, the need for software to lead to science which is verifiable and reproducible, policy issues such as how to measure software impact, choose licensing, or assign software credit, and the education and nurturing of early stage researchers in scientific software. 
Based on successful WSSSPE1-WSSSPE3 meetings at SC13, SC14, and NCAR in 2013, 2014, and 2015, the organizers expect over 60 submitted papers and around 75 attendees at the WSSSPE4 meeting that will be held immediately preceding the RSE (Research Software Engineer) Conference in Manchester, UK, in September 2016.  As with previous events, WSSSPE4 will provide an open and dynamic forum for the community to discuss experiences and approaches in the areas listed above. This proposal requests funds to support attendance at WSSSPE4 for US based researchers, particularly targeted at encouraging attendance by early stage researchers with attention to diversity.

@@

This is the fourth major event in the WSSSPE series. The first two WSSSPE workshops were held as part of SC in 2013 and 2014. These one-day workshops brought together researchers and practitioners to discuss challenges and issues around the development, deployment and maintenance of reusable software for academic research and education, and discussed potential solutions, moving from a presentation to a presentation and discussion format.  Reports from these workshops, and set of extended papers from them, have been published in special collections of the Journal of Open Research Software \cite{WSSSPE1-collection, WSSSPE2-collection, WSSSPE1, WSSSPE2}.
WSSSPE3, held at NCAR in 2015, grew out of the previous events, and focused almost entirely on facilitated discussions, with the aim of starting working groups to make changes in the practice of scientific software development leading to more sustainable software.  Two of these groups have continued
working together, leading to the FORCE11 Software Citation Principles and an in-progress paper
on best practices for sustainable scientific software development.  The report from WSSSPE3 has been submitted to the Journal of Open Research Software and is currently available as a preprint on arXiv~\cite{WSSSPE3}.

@@

\begin{quote}
{\bf Track 1 – Building a sustainable future for open-use research software} has the goals of defining a vision of the future of open-use research software, and in the workshop, initiating the activities that are needed to get there.  The idea of this track is to first think about where we want to be 5 to 10 years from now, without being too concerned with where we are today, and then to determine how we can move to this future.

{\bf Track 2 – Practices \& experiences in sustainable scientific software} has the goal of improving the quality of today’s research software and the experiences of its developers by sharing practices and experiences.  This track is focused on the current state of scientific software and what we can do to improve it in the short term, starting with where we are today.
\end{quote}

@@

The WSSSPE4 organizing committee comprises 
\begin{itemize}
\item Gabrielle Allen, University of Illinois Urbana-Champaign, USA
\item Jeffrey Carver, University of Alabama, USA
\item Sou-Cheng T. Choi, Illinois Institute of Technology, USA
\item Tom Crick, Cardiff Metropolitan University, UK
\item Michael R. Crusoe, Common Workflow Language project
\item Sandra Gesing, University of Notre Dame, USA
\item Robert Haines, University of Manchester, UK
\item Michael Heroux, Sandia National Laboratories, USA
\item Lorraine J. Hwang, University of California, Davis, USA
\item Daniel S. Katz, University of Illinois Urbana-Champaign, USA
\item Kyle E. Niemeyer, Oregon State University, USA
\item Manish Parashar, Rutgers University, USA
\item Colin C. Venters, University of Huddersfield, UK
\end{itemize}

The program committee currently has about 60 global members, from academia, industry, and national labs, with more invitations outstanding.  See the web site (\url{http://wssspe.researchcomputing.org.uk/wssspe4/}) for the specific names.

The tentative agenda for the workshop is

Sept 12 (pm) --- Joint session of both tracks

\begin{itemize}
\item Introduction
\item Keynote
\item Lightning talks
\item Updates on actions and activities from WSSSPE3 working groups
\item Discussion and planning for the remainder of WSSSPE4
\end{itemize}

Sept 13 (all day) and 14 (am) --- Parallel tracks

\begin{itemize}
\item Track 1: This will be a set of working sessions with a facilitated discussion, breakout sessions, report backs, and active writing towards the track goal of defining a vision of the future of open-use research software, and a plan of activities that are needed to get there.
\item Track 2: Presentations of position papers, experience papers, previously published works, and demos; and breakout sessions or unconference sessions.
\end{itemize}

Sept 14 (1:30 pm -- 5 pm) --- Joint session of both tracks

\begin{itemize}
\item Panel on best practices
\item Summary and discussion of each tracks’ progress
\item Planning for future events
\end{itemize}

@@

We are currently including a description of the opportunity for travel support in our call for papers and outreach material for the event to make sure that lack of funding is not a barrier to submitting a paper, and to make it clear that participation by early stage researchers is welcomed and encouraged.

@@

The WSSSPE4 organizing committee will review the applications and select the participants to be funded through the award.

Awarded funding is expected to be reimbursement of allowable expenses (travel, lodging, sustenance) of up to \$1,500 for participants traveling from outside Europe to the UK, up to \$1,000 for participants traveling from Europe outside the UK to the UK, and up to \$500 for participants traveling from within the UK.

@@

The National Center for Supercomputing Applications (NCSA) at the University of Illinois (U of I) will be the lead fiscal institution for this proposal to the Office of Advanced Cyberinfrastructure (ACI) at the National Science Foundation. NCSA staff will manage the travel for the participants and presenters funded through this proposal. US participants and presenters will arrange their flights through the travel agency of NCSA/University of Illinois, with reimbursed by NCSA/University of Illinois. The WSSSPE~2 organizing committee will review the applications and select the participants to be funded through this award.

@@

The focus of the WSSSPE4 workshop on sustainable software for science and engineering is a topic of very broad relevance and interest that cuts across many different disciplines and different categories of people from students, developers, faculty, etc. Areas of interest and focus in the workshop include education and career paths of those involved in the development of software.

@@

\section{Prior Support}

{\bf Daniel S. Katz}
was most recently supported by the NSF as a Program Officer under NSF ACI-1239751, made to the University of Chicago for \$999,966 for the period 3/26/2012 to 3/25/2016.  A large number of publications were produced under the PI's IRD time in this award, including the WSSSPE1 and WSSSPE2 reports~\cite{WSSSPE1, WSSSPE2}.
In addition, a number of NSF programs and official documents were produced under this award~\cite{NSF_software_vision}.

{\bf Gabrielle D. Allen}
\textcolor{red}{todo...}

{\bf Kyle E. Neimeyer}
\textcolor{red}{todo...}

@@

\bibliographystyle{abbrv}
\bibliography{wssspe}

@@

{\bf WSSSPE4}

{\bf Principal Investigators: Daniel Katz, Gabrielle Allen, Kyle Neimeyer}
